\blue{...Numerical investigation of viscous flow instabilities through heterogeneous porous media using adaptive and unstructured meshes under different mobility scenarios...}

### 2017-07-17 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
update

\medskip
Here, two numerical simulations were conducted with fixed and adaptive mesh resolutions to qualitatively assess model capability to capture fingers dynamics. In order to trigger the formation of fingers, regions with sharp synthetic permeability gradient were introduced. The computational space, Fig.~\ref{fig:testcase_heter_domain}a, consists of a rectangular domain fully saturated with a fluid (except by a squared region containing 50$\%$ of a second fluid -- wetting fluid phase, Fig.~\ref{fig:testcase_heter_domain}b) and divided into 5 regions with prescribed permeabilities. A no-flux boundary condition was imposed on upper and lower borders, whilst pure wetting fluid phase $\left(\text{\ie} S_{1}=1\right)$ is driven into the domain from the left-hand face at velocity $u_{1}=1$. Viscosity ratio of the two fluids is set to 10.

Figures \ref{fig:5regions_fixedmesh} and \ref{fig:5regions_adaptmesh} show fluid displacement in simulations performed with fixed and adaptive meshes. Both simulations started with a mesh of 13068 triangular (\PN[1]{2}) elements, however as flow dynamics evolve the number of elements of the simulation conducted with adaptive mesh oscillates from a minimum of 4400 to a maximum of 16430 (Fig.~\ref{fig:5regions_plottimenodeselements}). In both cases, fingers' development (formation, growth and coalescence) and fluid cross flows (through regions of distinct permeability at the top of the domain, $\mathbf{K}=\left\{2,3\right\}$) can be readily noticed.
The simulation performed with a fixed and relatively fine mesh
Sharp permeability gradient triggered the onset of fingers' formation, as shown in

\medskip
Having already verified that model is working for the hele-shaw experimental test-cases, now a more complicated case will be introduced with regions of different geometrical domains and different permeabilities. Initial and boundary conditions are described in fig.\ref{fig:testcase_heter_domain}, while inlet velocity is $1$ on the left hand side, and the pressure is set to $0$ on the outlet. What is expected is a significant growth of fingers when comparing mobility ratios that differ an order of magnitude. 
%(fig.\ref{fig:1testcase})
Figure \ref{fem_cv_represent_vr1_vr10} describes an initial comparison between the $VR=1$ and $VR=10$ case under the same domain, boundary and initial conditions and number of elements so as to underastand how the cross-flow in the upper layers can be developed compared with the fingers that will appear below. The purpose of these two cases was to verify that the model is working for cross flow cases thus will allow to proceed further and expand this technique into more complex formations. The next step is a comparison under the same mobility ratio of $10$, using the fixed and the adaptive mesh. For the fixed mesh we have $4600$ elements, but for the adaptive mesh we start with $400$ up to $600$, depends how fine we would like the mesh to be, and we reach up to $24000$. %Figure \ref{fig:testcase_heter_domain} gives a description of the domain that will be used.

\medskip
Figures \ref{fig:2testcase_a}%\ref{fig:2testcase_b}, \ref{fig:2testcase_c} 
-\ref{fig:2testcase_d} give a summary of the simulation for different time-steps as these are described. In this set, the interfacial instability is becoming more profound as the flow develops. Based on these two cases so far, an initial relation between the quantitative and qualitative outcome of our simulations can be derived. Fingers will develop faster in areas where permeabilities are higher. This can be seen from the overall sharp increase of velocity magnitude during the injection of the fluid $1$.

\subsection{ \blue{The concept of the preferential flow path. A 3D case demonstration} }\label{section:results_dinlet}

>>>>

\medskip
Here, two numerical simulations were conducted with fixed and adaptive mesh resolutions to qualitatively assess model capability to capture fingers dynamics. In order to trigger the formation of fingers, regions with sharp permeability gradient were introduced. The computational space, Fig.~\ref{fig:testcase_heter_domain}a, consists of a rectangular domain fully saturated with a fluid (except by a squared region containing 50$\%$ of a second fluid -- wetting fluid phase, Fig.~\ref{fig:testcase_heter_domain}b) and divided into 5 regions with prescribed permeability distribution. A no-flux boundary condition was imposed on upper and lower borders, whilst pure wetting fluid phase $\left(\text{\ie} S_{1}=1\right)$ is driven into the domain from the left-hand face at velocity $u_{1}=1$. Viscosity ratio was set to 10.

Figures \ref{fig:5regions_fixedmesh} and \ref{fig:5regions_adaptmesh} show fluid displacement in simulations performed with fixed and adaptive meshes. Both simulations started with a mesh of 13068 triangular (\PN[1]{2}) elements, however as flow dynamics evolve the number of elements of the simulation conducted with adaptive mesh oscillates from a minimum of 4400 to a maximum of 16430 (Fig.~\ref{fig:5regions_plottimenodeselements}). In both cases, fingers' development (formation, growth and coalescence) and fluid cross flows (through regions of distinct permeabilities at the top of the domain, $\mathbf{K}=\left\{2,3\right\}$) can be readily noticed. The simulation performed with a fixed and relatively fine mesh was able to capture the progressive development of fingers (as shown in Fig.~\ref{fig:5regions_AdaptZoom}b) prompted by sharp permeability gradient at the interface of regions with $\mathbf{K}=\left\{1,5\right\}$. In regions with no permeability gradient, formation and development of fingers (Fig.~\ref{fig:5regions_AdaptZoom}c) were not captured by the simulation conducted with fixed mesh. In both cases, dynamic adaptive mesh based on perturbation of phase saturations with imposed interpolation error estimate proved to be able to capture the onset of instabilities and fingers' development with relatively little computational overhead.

\subsection{3-D Channel Flows: Investigation of Preferential Flow Path and Formation of Fingers} \label{section:results_dinlet}

\blue{(a) Little intro of the case; (b) set up what are the main issues (theory) of the simulations you are showing in this section; (c) then briefly describe what are the strategies to tackle these issues; (d) then describe the numerical experiment w all BC and IC; (e) then finally describe the results with a clear conclusion based on observation but backed by theory and/or physics!!}
